---
title: "acorn_clean_notebook"
output: html_notebook
---
 

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lmtest)
library(tableone)
library(officer)

```

```{r}
library(readxl)
acorn_data <- read_excel("acorn_data.xlsx")
View(acorn_data)

overview <- acorn_data
```

```{r}
#look at missing data for white count

overview %>%
  sum(is.na(baseline_wbc))

#exclude 110 with missing values for baseline wbc

overview_new <- overview %>%
  filter(!is.na(baseline_wbc))

#look at distribution of white count for remaining patients
wbc_dist <- ggplot(overview_new, aes(x = baseline_wbc)) +
  geom_histogram()
#still some people with extremely high initial wbc

length(which(overview_new$baseline_wbc > 50)) #15 rows > 50
length(which(overview_new$baseline_wbc > 75)) #5 rows > 75
length(which(overview_new$baseline_wbc > 100)) #only 1 > 100, will remove that value

overview_new <- overview_new %>%
  filter(baseline_wbc < 100)

#remove outlier
wbc_dist_clean <- ggplot(overview_new, aes(x = baseline_wbc)) +
  geom_histogram() 

```


```{r}
#select variables of interest
smart_abx = overview_new%>%rename(id=record_id)%>%
                          mutate(CKD = ifelse(CKD==2,1,CKD))%>%
                          mutate(CKD = ifelse(is.na(CKD),0,CKD))%>%
                        dplyr::select(id, trt_group, inhosp_mortality28, death14_newrrt14, icufreedays28, delc14, comagdelcfreedays14, comagdelc14, persistent_renal_dysfunction28, rrtfreedays28, newrrt28, newrrt14, highest_cr_28, age, sex, race.eth, baseline_sofa, baseline_coma_delirium, incident_aki_with_onstudy_14.factor, CKD, sepsis, received_vancomycin_14, received_assigned_therapy_28, MAKE14, preill_creat_calculate, preill_creat, preill_creat_observed,
                                      baseline_vent, baseline_anyvasopressor, source_of_infection, location, priorhd, days_assigned_therapy_14, baseline_wbc, baseline_neutropenia)%>%
  mutate(n=1)%>%
  mutate(race_white = ifelse(race.eth=="White, Non-Hispanic", 1, 0))%>%
  mutate(coma= as.factor(comagdelcfreedays14))

smart_abx$incident_aki_with_onstudy_14.factor <- ordered(smart_abx$incident_aki_with_onstudy_14.factor,
                                                         levels = c("Alive without AKI", "Stage 1 AKI", 
                                     
                                                               "Stage 2 AKI", "Stage 3 AKI", "Died"))

smart_abx$source_of_infection <- as.factor(smart_abx$source_of_infection)

smart_abx$baseline_neutropnia <- as.factor(smart_abx$baseline_neutropenia)

```


```{r}
#look for missing data here
#count missing values in each variable

sapply(smart_abx, function(x) sum(is.na(x)))

#bar graph of missing values
freqDf <- data.frame(table(is.na(smart_abx))) 
  
# barplot for visualization 
barplot(freqDf$Freq , main = "Total Missing values", 
xlab = "Missing Data", ylab = "Frequency",  
        names.arg = c("FALSE","TRUE"), 
col = c("#80dfff","lightgreen")) 
  
# legend for barplot 
legend("topright", 
c("Non-Missing Values","Missing Values"), 
fill = c("#80dfff","lightgreen")) 
```


```{r}
#are missing variables values for wbc related to any predictors?

overview_missing <- overview_new %>%
  mutate(wbc_missing = ifelse(is.na(baseline_wbc), 0, 1))

missing_relationship <- glm(wbc_missing ~ age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + preill_creat + priorhd + source_of_infection + location, data = overview_missing, family = "binomial")
summary(missing_relationship)
```

```{r}
##creating Table 1
#characteristics to capture: age, sex, race, smoking, baseline creatinine, Charlson, baseline qSOFA, infection source

#table 1 with patient level stuff
myVars <- c("trt_group", "vancomycin_day0","received_vancomycin_14", "age",  "bmi",  "baseline_sofa", "daysaccurate_timezero_firstanyantibiotics", "CHARLSON", "sex", "race.eth", "baseline_vent", "prevalent_aki.factor", "sepsis", "source_of_infection", "baseline_anyvasopressor", "inhosp_mortality28", "MAKE14", "comagdelcfreedays14", "baseline_coma_delirium")

catVars <- c("trt_group", "vancomycin_day0","received_vancomycin_14","sex", "race.eth", "baseline_vent", "prevalent_aki.factor", "sepsis", "source_of_infection", "baseline_anyvasopressor", "inhosp_mortality28")

table1 <- CreateTableOne(vars = myVars, overview_new, factorVars = catVars)
table1

table1 <- print(table1, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
## Save to a CSV file
write.csv(table1, file = "myTable.csv")

#repeat, stratified by treatment group
table1_trt_group <- CreateTableOne(vars = myVars, overview_new, factorVars = catVars, strata = c("trt_group"))
table1_trt_group

```


```{r}
#interaction model
interaction_model <- function(predictor, data) {
modx <- glm(predictor ~ trt_group +  baseline_wbc + baseline_wbc*trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + location + preill_creat + priorhd, data = data, family = "binomial")
}

model_results <- function(result) {
  mod.sumx <- summary(result)
  mod.expx <- cbind(
    OR = exp(coef(result)),
    exp(confint(result)),
    p = mod.sumx$coefficients[, "Pr(>|z|)"]
  )
}

#test with 28-day mortality
mortality_int <- interaction_model(smart_abx$inhosp_mortality28, smart_abx)

#result
mortality_int_result <- model_results(mortality_int)

#no interaction model
no_interaction_model <- function(predictor, data) {
mody <- glm(predictor ~  trt_group + baseline_wbc + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + location + preill_creat + priorhd, data = data, family = "binomial")
}

```

```{r}
#effect modification by white count
#28 day mortality
moda <- glm(inhosp_mortality28 ~ trt_group +  baseline_wbc + baseline_wbc*trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + location + preill_creat + priorhd, data = smart_abx, family = "binomial")

moda_results <- model_results(moda)
moda_results
 
modb <- glm(inhosp_mortality28 ~  trt_group + baseline_wbc + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + location + preill_creat + priorhd, data = smart_abx, family = "binomial")
  mod.sumb <- summary(modb)
  mod.expb <- cbind(
    OR = exp(coef(modb)),
    exp(confint(modb)),
    p = mod.sumb$coefficients[, "Pr(>|z|)"]
  )
  
  lrtest(moda, modb)
```


```{r}
#make 14
make14_int <- interaction_model(smart_abx$MAKE14, smart_abx)

make14 <- no_interaction_model(smart_abx$MAKE14, smart_abx)

#interaction results
make14_int_result <- model_results(make14_int)

#no interaction results
make14_results <- model_results(make14)

lrtest(make14_int, make14)
```

```{r}
#incident aki
aki_int = MASS::polr(incident_aki_with_onstudy_14.factor ~ trt_group +  baseline_wbc + baseline_wbc*trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + preill_creat + priorhd + source_of_infection + location, data = smart_abx, Hess = TRUE)
aki = MASS::polr(incident_aki_with_onstudy_14.factor ~ trt_group +  baseline_wbc + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + preill_creat + priorhd + source_of_infection + location, data = smart_abx, Hess = TRUE)

anova(aki_int, aki, test ="Chisq")

#interaction results
aki_int_sum <- summary(aki_int)
  
aki_int_ci <- confint(aki_int)
aki_int_results <- exp(cbind(OR = coef(aki_int), aki_int_ci))

#no interaction results
aki_sum <- summary(aki)

aki_ci <- confint(aki)
aki_results <- exp(cbind(OR = coef(aki), aki_ci))

#detailed interaction results with p values
summary_aki_cont <- summary(aki_int)

# Extract coefficients and standard errors
coefficients <- coef(summary_aki_cont)
z_values <- coefficients[, "Value"] / coefficients[, "Std. Error"]  # Calculate Z-values
p_values <- 2 * pnorm(abs(z_values), lower.tail = FALSE)  # Two-tailed p-values

# Combine results into a data frame
aki_cont_results <- data.frame(
  Coefficient = coefficients[, "Value"],
  `Std. Error` = coefficients[, "Std. Error"],
  `Z-value` = z_values,
  `P-value` = p_values
)
print(aki_cont_results)

# Calculate odds ratios and confidence intervals
conf_intervals <- confint(aki_int)  # Profile likelihood confidence intervals
odds_ratios <- exp(coef(aki_int))  # Exponentiate coefficients
conf_intervals_exp <- exp(conf_intervals)  # Exponentiate confidence intervals

# Combine ORs and CI into a data frame
aki_or_results <- data.frame(
  OR = odds_ratios,
  `2.5% CI` = conf_intervals_exp[, 1],
  `97.5% CI` = conf_intervals_exp[, 2]
)
print(aki_or_results)

```

```{r}
#incidence of coma or delirium
ever_coma_int <- interaction_model(smart_abx$comagdelc14, smart_abx)

ever_coma <- no_interaction_model(smart_abx$comagdelc14, smart_abx)

lrtest(ever_coma_int, ever_coma)

#interaction results
  
coma_int_results <- model_results(ever_coma_int)

#no interaction results
coma_results <- model_results(ever_coma)

```


```{r}
#sensitivity analysis: sepsis or vancomycin only
sepsis_vanc <- smart_abx %>%
  filter(received_vancomycin_14 == 1 & sepsis == "Yes")

s_v_mortality_int <- interaction_model(sepsis_vanc$inhosp_mortality28, sepsis_vanc)

s_v_mortality <- no_interaction_model(sepsis_vanc$inhosp_mortality28, sepsis_vanc)

lrtest(s_v_mortality_int, s_v_mortality)

#mortality results
s_v_mortality_int_results <- model_results(s_v_mortality_int)

s_v_mortality_results <- model_results(s_v_mortality)
  
```




```{r}
#sensitivity: vanc only
vanc <- smart_abx %>%
  filter(received_vancomycin_14 == 1)

v_mortality_int <- interaction_model(vanc$inhosp_mortality28, vanc)

v_mortality <- no_interaction_model(vanc$inhosp_mortality28, vanc)

lrtest(v_mortality_int, v_mortality)

#mortality results
v_mortality_int_results <- model_results(v_mortality_int)

v_mortality_results <- model_results(v_mortality)
```
```{r}
#sensitivity: baseline sepsis
sepsis <- smart_abx %>%
  filter(sepsis == "Yes")

s_mortality_int <- interaction_model(sepsis$inhosp_mortality28, sepsis)

s_mortality <- no_interaction_model(sepsis$inhosp_mortality28, sepsis)

lrtest(s_mortality_int, s_mortality)

#mortality results
s_mortality_int_results <- model_results(s_mortality_int)

s_mortality_results <- model_results(s_mortality)
```



```{r}
#sensitivity: exclude abdominal infection
no_abdominal <- smart_abx %>%
  filter(source_of_infection != "Intra-abdominal")

no_abdominal$source_of_infection <- droplevels(no_abdominal$source_of_infection)

no_abdominal_mortality_int <- interaction_model(no_abdominal$inhosp_mortality28, no_abdominal)

no_abdominal_mortality <- no_interaction_model(no_abdominal$inhosp_mortality28, no_abdominal)

lrtest(no_abdominal_mortality_int, no_abdominal_mortality)

#mortality results
no_abdominal_mortality_int_results <- model_results(no_abdominal_mortality_int)

no_abdominal_mortality_results <- model_results(no_abdominal_mortality)
```
```{r}
#fast way to get results onto table
install.packages("jtools")
require(jtools)
install.packages("flextable")
require(flextable)
install.packages("openxlsx")
require(openxlsx)

coma_combined_results <- export_summs(ever_coma_int, ever_coma, exp = TRUE,
             error_format = "({conf.low}, {conf.high})", to.file = "xlsx", file.name = "test.xlsx")
```

```{r}
#data visualizaiton by quartile
#create quartiles of initial wbc
smart_abx <- smart_abx %>%
  mutate(wbc_quartile = ntile(baseline_wbc, 4))

#make labeled factor for plots
overview_new <- overview_new %>%
  mutate(wbc_quintile = recode_factor(ntile(baseline_wbc, 4), `1` = "<7", `2` = "7-11", `3` = "11-16", `4` = ">16"))

```

```{r}
#table one by quartile
#table 1 with patient level stuff
myVars <- c("trt_group", "vancomycin_day0","received_vancomycin_14", "age",  "bmi",  "baseline_sofa", "daysaccurate_timezero_firstanyantibiotics", "CHARLSON", "sex", "race.eth", "baseline_vent", "prevalent_aki.factor", "sepsis", "source_of_infection", "baseline_anyvasopressor", "inhosp_mortality28", "MAKE14", "comagdelcfreedays14", "baseline_coma_delirium")

catVars <- c("trt_group", "vancomycin_day0","received_vancomycin_14","sex", "race.eth", "baseline_vent", "prevalent_aki.factor", "sepsis", "source_of_infection", "baseline_anyvasopressor", "inhosp_mortality28")

table1 <- CreateTableOne(vars = myVars, overview_new, factorVars = catVars, strata = c("wbc_quartile"))
table1

table1 <- print(table1, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
## Save to a CSV file
write.csv(table1, file = "myTable.csv")
```

```{r}
#look at baseline characteristics only in highest quartile

overview_new <- overview_new %>%
  mutate(wbc_quartile = ntile(baseline_wbc, 4))

#make labeled factor for plots
overview_new <- overview_new %>%
  mutate(wbc_quintile = recode_factor(ntile(baseline_wbc, 4), `1` = "<7", `2` = "7-11", `3` = "11-16", `4` = ">16"))

overview_q4 <- subset(overview_new, wbc_quartile == 4)


#table 1 with patient level stuff
Vars <- c("trt_group", "vancomycin_day0","received_vancomycin_14", "age",  "bmi",  "baseline_sofa", "daysaccurate_timezero_firstanyantibiotics", "CHARLSON", "sex", "race.eth", "baseline_vent", "prevalent_aki.factor", "sepsis", "source_of_infection", "baseline_anyvasopressor", "inhosp_mortality28", "MAKE14", "comagdelcfreedays14", "baseline_coma_delirium", "baseline_neutropenia", "ever_transplant")

catVars <- c("vancomycin_day0","received_vancomycin_14","sex", "race.eth", "baseline_vent", "prevalent_aki.factor", "sepsis", "source_of_infection", "baseline_anyvasopressor")

q4_table1 <- CreateTableOne(vars = Vars, overview_q4, factorVars = catVars, strata = c("trt_group"))
q4_table1

q4_table1 <- print(q4_table1, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
## Save to a CSV file
write.csv(q4_table1, file = "q4Table.csv")

#repeat for each quartile
#q1
overview_q1 <- subset(overview_new, wbc_quintile == "<7")
q1_table <- CreateTableOne(vars = Vars, overview_q1, factorVars = catVars, strata = c("trt_group"))

#q2
overview_q2 <- subset(overview_new, wbc_quintile == "7-11")
q2_table <- CreateTableOne(vars = Vars, overview_q2, factorVars = catVars, strata = c("trt_group"))

q2_table <- print(q2_table, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
## Save to a CSV file
write.csv(q2_table, file = "q2_Table.csv")

#q3
overview_q3 <- subset(overview_new, wbc_quintile == "11-16")
q3_table <- CreateTableOne(vars = Vars, overview_q3, factorVars = catVars, strata = c("trt_group"))

q3_table <- print(q3_table, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
## Save to a CSV file
write.csv(q3_table, file = "q3_Table.csv")
```

```{r}
#28 day mortality by quartile
moda <- glm(inhosp_mortality28 ~ trt_group +  factor(wbc_quartile) + factor(wbc_quartile)*trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + factor(source_of_infection) + location + preill_creat_calculate + priorhd, data = smart_abx, family = "binomial")
  mod.suma <- summary(moda)
  mod.expa <- cbind(
    OR = exp(coef(moda)),
    exp(confint(moda)),
    p = mod.suma$coefficients[, "Pr(>|z|)"]
  )
 
  modb <- glm(inhosp_mortality28 ~  trt_group + factor(wbc_quartile) + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + factor(source_of_infection) + location + preill_creat_calculate + priorhd, data = smart_abx, family = "binomial")
  mod.sumb <- summary(modb)
  mod.expb <- cbind(
    OR = exp(coef(modb)),
    exp(confint(modb)),
    p = mod.sumb$coefficients[, "Pr(>|z|)"]
  )
  
  lrtest(moda, modb)
  
```

```{r}
#make14 by quartile
make_quart_int <- glm(MAKE14 ~ trt_group +  factor(wbc_quartile) + factor(wbc_quartile)*trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + factor(source_of_infection) + location + preill_creat + priorhd, data = smart_abx, family = "binomial")
  mod.sum.make.int <- summary(make_quart_int)
  make_quart_int_results <- cbind(
    OR = exp(coef(make_quart_int)),
    exp(confint(make_quart_int)),
    p = mod.sum.make.int$coefficients[, "Pr(>|z|)"]
  )
 
  make_quart <- glm(MAKE14 ~  trt_group + wbc_quartile + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + factor(source_of_infection) + location + preill_creat + priorhd, data = smart_abx, family = "binomial")
  mod.sum.make <- summary(make_quart)
  make_quart_results <- cbind(
    OR = exp(coef(make_quart)),
    exp(confint(make_quart)),
    p = mod.sum.make$coefficients[, "Pr(>|z|)"]
  )
  
  lrtest(make_quart_int, make_quart)
```

```{r}

smart_abx$wbc_quartile <- as.factor(smart_abx$wbc_quartile)

#incident aki
aki.a <- MASS::polr(incident_aki_with_onstudy_14.factor ~ trt_group + factor(wbc_quartile) + factor(wbc_quartile)*trt_group + age + sex +  baseline_sofa + baseline_vent + baseline_anyvasopressor + preill_creat + factor(source_of_infection) + location, data = smart_abx, Hess = TRUE)

aki_quart_results <- exp(cbind(OR = coef(aki.a), (confint(aki.a))))

aki.b <- MASS::polr(incident_aki_with_onstudy_14.factor ~ trt_group + factor(wbc_quartile) + age + sex +  baseline_sofa + baseline_vent + baseline_anyvasopressor + preill_creat + factor(source_of_infection) + location, data = smart_abx, Hess = TRUE)



aki_quart_results <- exp(cbind(OR = coef(aki.a), exp(conf_intervals)))

summary_aki <- summary(aki.a)

# Extract coefficients and standard errors
coefficients <- coef(summary_aki)
z_values <- coefficients[, "Value"] / coefficients[, "Std. Error"]  # Calculate Z-values
p_values <- 2 * pnorm(abs(z_values), lower.tail = FALSE)  # Two-tailed p-values

# Combine results into a data frame
results <- data.frame(
  Coefficient = coefficients[, "Value"],
  `Std. Error` = coefficients[, "Std. Error"],
  `Z-value` = z_values,
  `P-value` = p_values
)
print(results)

# Calculate odds ratios and confidence intervals
conf_intervals <- confint(aki.a)  # Profile likelihood confidence intervals
odds_ratios <- exp(coef(aki.a))  # Exponentiate coefficients
conf_intervals_exp <- exp(conf_intervals)  # Exponentiate confidence intervals

# Combine ORs and CI into a data frame
or_results <- data.frame(
  OR = odds_ratios,
  `2.5% CI` = conf_intervals_exp[, 1],
  `97.5% CI` = conf_intervals_exp[, 2]
)
print(or_results)

```

```{r}
#coma or delirium incidence
ever_coma_int_quart = glm(comagdelc14 ~ trt_group +  factor(wbc_quartile) + factor(wbc_quartile)*trt_group + age + sex+  baseline_sofa  + baseline_anyvasopressor + preill_creat + priorhd + factor(source_of_infection) + location, data = smart_abx, family = "binomial")

coma_quart_results <- exp(cbind(OR = coef(ever_coma_int_quart), confint(ever_coma_int_quart)))

ever_coma_quart = glm(comagdelc14 ~ trt_group +  wbc_quartile + age + sex+  baseline_sofa  + baseline_anyvasopressor + preill_creat + priorhd + factor(source_of_infection) + location, data = smart_abx, family = "binomial")

lrtest(ever_coma_int_quart, ever_coma_quart)

ever_coma_int_quart_sum <- summary(ever_coma_int_quart)

ever_coma_quart_int_results <- cbind(
    OR = exp(coef(ever_coma_int_quart)),
    exp(confint(ever_coma_int_quart)),
    p = ever_coma_int_quart_sum$coefficients[, "Pr(>|z|)"]
  )

ever_coma_quart_sum <- summary(ever_coma_quart)
ever_coma_quart_results <- cbind(
    OR = exp(coef(ever_coma_quart)),
    exp(confint(ever_coma_quart)),
    p = ever_coma_quart_sum$coefficients[, "Pr(>|z|)"]
  )
```


```{r}
#odds of outcome in each quartile
#mortality
stratified_model <- function(outcome, data) {
    modx <- glm(outcome ~ trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + factor(source_of_infection) + location + preill_creat + priorhd, data = data, family = "binomial")
  mod.sumx <- summary(modx)
  OR.quartile <- cbind(
    OR = exp(coef(modx)),
    exp(confint(modx)),
    p = mod.sumx$coefficients[, "Pr(>|z|)"]
  )
  return(OR.quartile)
}

#mortality in each quartile
q1_OR <- stratified_model(q1$inhosp_mortality28, q1)
q2_OR <- stratified_model(q2$inhosp_mortality28, q2)
q3_OR <- stratified_model(q3$inhosp_mortality28, q3)
q4_OR <- stratified_model(q4$inhosp_mortality28, q4)

#install package to make forest plot
install.packages("forestploter")
library(forestploter)

quartile_OR_data <- tibble('WBC Quartile' = c("<7", "7-11", "11-16", ">16"), 
                           OR = c(1.68, 1.49, 0.92, 0.51), 
                           LL = c(0.84, 0.64, 0.43, 0.29), 
                           UL = c(3.50, 3.52, 1.94, 0.90))

# Add a blank column for the forest plot to display CI.
# Adjust the column width with space, and increase the number of spaces below 
# to have a larger area to draw the CI. 
quartile_OR_data$` ` <- paste(rep(" ", 35), collapse = " ")


# Create a confidence interval column to display
quartile_OR_data$`OR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     quartile_OR_data$OR, quartile_OR_data$LL, quartile_OR_data$UL)

tm <- forest_theme(core=list(fg_params=list(hjust=01, x=0.9)), rowhead=list(fg_params=list(hjust=0.5, x=0.5)), padding = unit(c(50, 50), "mm"))

OR_forest <- forest(quartile_OR_data[, c(1, 5:6)],
            est = quartile_OR_data$OR,
            lower = quartile_OR_data$LL, 
            upper = quartile_OR_data$UL,
            ci_column = 2,
            ref_line = 1,
            xlim = c(0, 2),
            ticks_at = c(0.5, 1, 1.5, 2),
            title = "28-Day Mortality", 
            theme = tm)
OR_forest
```


```{r}
#make14
#make14
q1_make_OR <- stratified_model(q1$MAKE14, q1)
q2_make_OR <- stratified_model(q2$MAKE14, q2)
q3_make_OR <- stratified_model(q3$MAKE14, q3)
q4_make_OR <- stratified_model(q4$MAKE14, q4)

quartile_OR_data <- tibble('WBC Quartile' = c("<7", "7-11", "11-16", ">16"), 
                           OR = c(1.38, 0.84, 0.89, 0.57), 
                           LL = c(0.71, 0.39, 0.44, 0.33), 
                           UL = c(2.76, 1.77, 1.81, 0.96))

# Add a blank column for the forest plot to display CI.
# Adjust the column width with space, and increase the number of spaces below 
# to have a larger area to draw the CI. 
quartile_OR_data$` ` <- paste(rep(" ", 35), collapse = " ")


# Create a confidence interval column to display
quartile_OR_data$`OR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     quartile_OR_data$OR, quartile_OR_data$LL, quartile_OR_data$UL)

tm <- forest_theme(core=list(fg_params=list(hjust=01, x=0.9)), rowhead=list(fg_params=list(hjust=0.5, x=0.5)), padding = unit(c(50, 50), "mm"))

make_OR_forest <- forest(quartile_OR_data[, c(1, 5:6)],
            est = quartile_OR_data$OR,
            lower = quartile_OR_data$LL, 
            upper = quartile_OR_data$UL,
            ci_column = 2,
            ref_line = 1,
            xlim = c(0, 2),
            ticks_at = c(0.5, 1, 1.5, 2),
            title = "MAKE14", 
            theme = tm)
make_OR_forest
```


```{r}
#incident aki
stratified_aki_function <- function(data) {
  model <- MASS::polr(incident_aki_with_onstudy_14.factor ~ trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + preill_creat + priorhd + factor(source_of_infection) + location, data = data, Hess = TRUE)

#results table
ci <- confint.default(model)
results <- exp(cbind(OR = coef(model), ci))

return(results)
}


q1_aki_or <- stratified_aki_function(q1)
q2_aki_or <- stratified_aki_function(q2)
q3_aki_or <- stratified_aki_function(q3)
q4_aki_or <- stratified_aki_function(q4)

quartile_OR_data <- tibble('WBC Quartile' = c("<7", "7-11", "11-16", ">16"), 
                           OR = c(1.16, 1.27, 1.10, 0.90), 
                           LL = c(0.78, 0.82, 0.72, 0.64), 
                           UL = c(1.73, 1.96, 1.70, 1.28))

# Add a blank column for the forest plot to display CI.
# Adjust the column width with space, and increase the number of spaces below 
# to have a larger area to draw the CI. 
quartile_OR_data$` ` <- paste(rep(" ", 35), collapse = " ")


# Create a confidence interval column to display
quartile_OR_data$`OR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     quartile_OR_data$OR, quartile_OR_data$LL, quartile_OR_data$UL)

tm <- forest_theme(core=list(fg_params=list(hjust=01, x=0.9)), rowhead=list(fg_params=list(hjust=0.5, x=0.5)), padding = unit(c(50, 50), "mm"))

aki_OR_forest <- forest(quartile_OR_data[, c(1, 5:6)],
            est = quartile_OR_data$OR,
            lower = quartile_OR_data$LL, 
            upper = quartile_OR_data$UL,
            ci_column = 2,
            ref_line = 1,
            xlim = c(0, 2),
            ticks_at = c(0.5, 1, 1.5, 2),
            title = "Incident AKI", 
            theme = tm)
aki_OR_forest


```

```{r}
#coma delirium
#ever coma by quartile to add to figure
#ever coma in each quartile
q1_coma_OR <- stratified_model(q1$comagdelc14, q1)
q2_coma_OR <- stratified_model(q2$comagdelc14, q2)
q3_coma_OR <- stratified_model(q3$comagdelc14, q3)
q4_coma_OR <- stratified_model(q4$comagdelc14, q4)

#aki forest
quartile_OR_data <- tibble('WBC Quartile' = c("<7", "7-11", "11-16", ">16"), 
                           OR = c(0.82, 0.79, 0.58, 1.12), 
                           LL = c(0.44, 0.39, 0.32, 0.67), 
                           UL = c(1.50, 1.56, 1.02, 1.88))

# Add a blank column for the forest plot to display CI.
# Adjust the column width with space, and increase the number of spaces below 
# to have a larger area to draw the CI. 
quartile_OR_data$` ` <- paste(rep(" ", 35), collapse = " ")


# Create a confidence interval column to display
quartile_OR_data$`OR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     quartile_OR_data$OR, quartile_OR_data$LL, quartile_OR_data$UL)


tm <- forest_theme(core=list(fg_params=list(hjust=01, x=0.9)), rowhead=list(fg_params=list(hjust=0.5, x=0.5)), padding = unit(c(50, 50), "mm"))

coma_OR_forest <- forest(quartile_OR_data[, c(1, 5:6)],
            est = quartile_OR_data$OR,
            lower = quartile_OR_data$LL, 
            upper = quartile_OR_data$UL,
            ci_column = 2,
            ref_line = 1,
            arrow_lab = c("Favors piperacillin/tazobactam", "Favors Cefepime"),
            xlim = c(0, 2),
            ticks_at = c(0.5, 1, 1.5, 2),
            title = "Coma or Delirium", 
            theme = tm)
coma_OR_forest
```


```{r}
#powerful way to get data by quartile that involves no typing!
#could also use lapply instead of map or vice versa

library(broom)
library(purrr)
library(forcats)

quartile_OR_data <- smart_abx %>%
  group_by(wbc_quartile) %>%
  nest() %>%
  mutate(
    fit = map(data, function(df) glm(inhosp_mortality28 ~ trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + preill_creat + priorhd, data = df, family = "binomial")),
    output = map(fit, tidy),
    ci = map(fit, confint)) %>%
  unnest(c(output, ci)) %>%
  filter(term == "trt_groupzosyn") %>% 
  mutate(
    OR = exp(estimate),
    LL = exp(ci[,1]),
    UL = exp(ci[2])
  ) %>% 
  select(wbc_quartile, OR, LL, UL)  %>%
  arrange(wbc_quartile) %>%
  mutate('WBC Quartile' = factor(wbc_quartile, levels = c(1, 2, 3, 4), labels = c("<7", "7-11", "11-16", ">16"))) %>%
  select('WBC Quartile', OR, LL, UL)
```


```{r}
#calculate marginal effects of model to estimate effect size

install.packages("marginaleffects")
library(marginaleffects)

# Plot conditional adjusted predictions
predictions_plot <- plot_predictions(moda, condition = list("baseline_wbc", "trt_group"))

#plot conditional marginal effects--how the derivative of hospital mortality with respect to treatment group varies as a function of baseline wbc 
plot_slopes(moda, variables = "trt_group", condition = list("baseline_wbc"))

#obtain average slopes (marginal effects over the whole model)

avg_slopes(moda, variables = "baseline_wbc", by = "trt_group")

#average comparisons
avg_comparisons(moda, variables = "baseline_wbc", by = "trt_group")
```
```{r}
#trying some things by quartile
mod_quart <- glm(inhosp_mortality28 ~ trt_group +  wbc_quartile + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + factor(source_of_infection) + location + preill_creat_calculate + priorhd, data = smart_abx, family = "binomial")

#estimate the partial derivative of in hospital mortality with respect to wbc quartile by treatment group
slopes(
  mod_quart,
  variables = "wbc_quartile",
  newdata = datagrid("trt_group"))

#make a tree
install.packages("partykit")
library(partykit)

tree <- ctree(
    inhosp_mortality28 ~ wbc_quartile + as.factor(trt_group),
    data = smart_abx,
    control = ctree_control(maxdepth = 2)
)

plot(tree)

dat <- transform(smart_abx, nodeid = predict(tree, type = "node"))
comparisons(mod_quart,
    newdata = dat,
    by = "nodeid")
```

```{r}
#wbc quartile characteristics
smart_abx %>%
  group_by(wbc_quartile) %>%
  summarize(
    mean = mean(baseline_wbc),
    range = range(baseline_wbc))
  
```

## Revisions

```{r}
# look at different WBC cutoffs

#create quartiles of initial wbc
smart_abx <- smart_abx %>%
  mutate(wbc_clinical = case_when(
    baseline_wbc <= 4 ~ "<4",
    baseline_wbc > 4 & baseline_wbc <= 11 ~ "4-11",
    baseline_wbc > 11 ~ ">11"
  ))

smart_abx %>%
  group_by(wbc_clinical) %>%
  count()

smart_abx <- smart_abx %>%
  mutate(wbc_clinical = factor(wbc_clinical, levels = c("<4", "4-11", ">11")))
  



```


```{r}
#plot for new wbc levels

smart_abx %>%
  group_by(wbc_clinical, trt_group) %>%
  summarise(
    N = sum(n),
    death = mean(inhosp_mortality28),
    se = sd(inhosp_mortality28)/sqrt(N), .groups = 'drop') %>%
  ggplot(aes(x = wbc_clinical, y = death, color = trt_group)) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin = death - se, ymax = death + se), width=0.2) +  # Add error bars
  labs(
    title = "Mortality rate by WBC Clinical Cutoffs and Treatment Group",
    x = "WBC Clinical Cutoffs",
    y = "Mortality Rate"
  ) +
  theme_classic()
  
```

```{r}
#modeling by wbc clinical cutoffs

library(broom)
library(purrr)
library(forcats)

clinical_OR_data <- smart_abx %>%
  group_by(wbc_clinical) %>%
  nest() %>%
  mutate(
    fit = map(data, function(df) glm(inhosp_mortality28 ~ trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + preill_creat + priorhd, data = df, family = "binomial")),
    output = map(fit, tidy),
    ci = map(fit, confint)) %>%
  unnest(c(output, ci)) %>%
  filter(term == "trt_groupzosyn") %>% 
  mutate(
    OR = exp(estimate),
    LL = exp(ci[,1]),
    UL = exp(ci[2])
  ) %>% 
  select(wbc_clinical, OR, LL, UL)  %>%
  arrange(wbc_clinical) 

clinical_OR_data$` ` <- paste(rep(" ", 25), collapse = " ")


# Create a confidence interval column to display
clinical_OR_data$`OR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     clinical_OR_data$OR, clinical_OR_data$LL, clinical_OR_data$UL)


tm <- forest_theme(core=list(fg_params=list(hjust=01, x=0.9)), rowhead=list(fg_params=list(hjust=0.5, x=0.5)), padding = unit(c(50, 50), "mm"))

clinical_OR_forest <- forest(clinical_OR_data[, c(1, 5:6)],
            est = clinical_OR_data$OR,
            lower = clinical_OR_data$LL, 
            upper = clinical_OR_data$UL,
            ci_column = 2,
            ref_line = 1,
            arrow_lab = c("Favors piperacillin/tazobactam", "Favors Cefepime"),
            xlim = c(0, 3),
            ticks_at = c(0.5, 1, 1.5, 2, 2.5),
            title = "Mortality with Clinical Cutoff", 
            theme = tm)
clinical_OR_forest
```



```{r}
#wbc cutoffs using sirs cutoffs

smart_abx <- smart_abx %>%
  mutate(wbc_sirs = case_when(
    baseline_wbc <= 4 ~ "<4",
    baseline_wbc > 4 & baseline_wbc <= 12 ~ "4-12",
    baseline_wbc > 12 ~ ">12"
  ))

smart_abx %>%
  group_by(wbc_sirs) %>%
  count()

smart_abx <- smart_abx %>%
  mutate(wbc_sirs = factor(wbc_sirs, levels = c("<4", "4-12", ">12")))
  



```
```{r}
# data visualization using sirs cutoffs
smart_abx %>%
  group_by(wbc_sirs, trt_group) %>%
  summarise(
    N = sum(n),
    death = mean(inhosp_mortality28),
    se = sd(inhosp_mortality28)/sqrt(N), .groups = 'drop') %>%
  ggplot(aes(x = wbc_sirs, y = death, color = trt_group)) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin = death - se, ymax = death + se), width=0.2) +  # Add error bars
  labs(
    title = "Mortality rate by WBC SIRS Cutoffs and Treatment Group",
    x = "WBC SIRS Groups",
    y = "Mortality Rate"
  ) +
  theme_classic()
```
```{r}
# modeling using sirs cutoffs
sirs_OR_data <- smart_abx %>%
  group_by(wbc_sirs) %>%
  nest() %>%
  mutate(
    fit = map(data, function(df) glm(inhosp_mortality28 ~ trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + preill_creat + priorhd, data = df, family = "binomial")),
    output = map(fit, tidy),
    ci = map(fit, confint)) %>%
  unnest(c(output, ci)) %>%
  filter(term == "trt_groupzosyn") %>% 
  mutate(
    OR = exp(estimate),
    LL = exp(ci[,1]),
    UL = exp(ci[2])
  ) %>% 
  select(wbc_sirs, OR, LL, UL)  %>%
  arrange(wbc_sirs) 

sirs_OR_data$` ` <- paste(rep(" ", 35), collapse = " ")


# Create a confidence interval column to display
sirs_OR_data$`OR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     sirs_OR_data$OR, sirs_OR_data$LL, sirs_OR_data$UL)


tm <- forest_theme(core=list(fg_params=list(hjust=01, x=0.9)), rowhead=list(fg_params=list(hjust=0.5, x=0.5)), padding = unit(c(50, 50), "mm"))

sirs_OR_forest <- forest(sirs_OR_data[, c(1, 5:6)],
            est = sirs_OR_data$OR,
            lower = sirs_OR_data$LL, 
            upper = sirs_OR_data$UL,
            ci_column = 2,
            ref_line = 1,
            arrow_lab = c("Favors piperacillin/tazobactam", "Favors Cefepime"),
            xlim = c(0, 3),
            ticks_at = c(0.5, 1, 1.5, 2, 2.5),
            title = "Mortality with SIRS Cutoffs", 
            theme = tm)
sirs_OR_forest
```
```{r}
# wbc epic

smart_abx <- smart_abx %>%
  mutate(wbc_epic = case_when(
    baseline_wbc <= 4 ~ "<4",
    baseline_wbc > 4 & baseline_wbc <= 10 ~ "4-10",
    baseline_wbc > 10 ~ ">10"
  ))

smart_abx %>%
  group_by(wbc_epic) %>%
  count()

smart_abx <- smart_abx %>%
  mutate(wbc_epic = factor(wbc_epic, levels = c("<4", "4-10", ">10")))
  
```

```{r}
# wbc epic modeling

epic_OR_data <- smart_abx %>%
  mutate(`WBC Epic Cutoffs` = wbc_epic) %>%
  group_by(`WBC Epic Cutoffs`) %>%
  nest() %>%
  mutate(
    fit = map(data, function(df) glm(inhosp_mortality28 ~ trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + preill_creat + priorhd, data = df, family = "binomial")),
    output = map(fit, tidy),
    ci = map(fit, confint)) %>%
  unnest(c(output, ci)) %>%
  filter(term == "trt_groupzosyn") %>% 
  mutate(
    OR = exp(estimate),
    LL = exp(ci[,1]),
    UL = exp(ci[2])
  ) %>% 
  select(`WBC Epic Cutoffs`, OR, LL, UL)  %>%
  arrange(`WBC Epic Cutoffs`) 

epic_OR_data$` ` <- paste(rep(" ", 35), collapse = " ")


# Create a confidence interval column to display
epic_OR_data$`OR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     epic_OR_data$OR, epic_OR_data$LL, epic_OR_data$UL)


tm <- forest_theme(base_size = 10, core=list(fg_params=list(hjust=01, x=0.9)), rowhead=list(fg_params=list(hjust=0.5, x=1)), padding = unit(c(50, 80, 50, 50), "mm"))

epic_OR_forest <- forest(epic_OR_data[, c(1, 5:6)],
            est = epic_OR_data$OR,
            lower = epic_OR_data$LL, 
            upper = epic_OR_data$UL,
            ci_column = 2,
            ref_line = 1,
            arrow_lab = c("Favors piperacillin/tazobactam", "Favors Cefepime"),
            xlim = c(0, 3),
            ticks_at = c(0.5, 1, 1.5, 2, 2.5),
            title = "Figure E2: Mortality with Epic Cutoffs", 
            theme = tm)
epic_OR_forest

```

## look at neutropenia

looks like in general wbc is low in this cohort though there are a few with higher wbc which seems potentially not correct
```{r}
#neutropenia

overview_new %>%
  group_by(baseline_neutropenia) %>%
  count()

#wbc distribution

overview_new %>%
  filter(baseline_neutropenia == "Yes") %>%
  ggplot(aes(x= baseline_wbc)) +
  geom_histogram()

overview_new %>%
  filter(baseline_neutropenia == "Yes") %>%
  group_by(trt_group) %>%
  mutate(N = n(),
         death = mean(inhosp_mortality28),
         se = sd(inhosp_mortality28)/sqrt(N), .groups = 'drop') %>%
  ggplot(aes(x = trt_group, y = death)) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin = death - se, ymax = death + se), width=0.2) +  # Add error bars
  labs(
    title = "Mortality rate by Treatment group among neutropenic patients",
    x = "Treatment group",
    y = "Mortality Rate"
  ) +
  theme_classic()
```

```{r}
# exclude neutropenic patients

no_neutropenia <- smart_abx %>%
  filter(baseline_neutropenia == "No")

neutropenia_excluded <- interaction_model(no_neutropenia$inhosp_mortality28, no_neutropenia)

results <- model_results(neutropenia_excluded)
results

no_int_no_neutro <- no_interaction_model(no_neutropenia$inhosp_mortality28, no_neutropenia)

lrtest(neutropenia_excluded, no_int_no_neutro)

# add neutropenia to existing model

neutropenia_included <- glm(inhosp_mortality28 ~ trt_group +  baseline_wbc + baseline_wbc*trt_group + baseline_neutropenia + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + location + preill_creat + priorhd, data = smart_abx, family = "binomial")

new_results <- model_results(neutropenia_included)
new_results
```

### transplant

```{r}
#transplant

overview_new %>%
  group_by(transplant) %>%
  count()

overview_new <- overview_new %>%
  mutate(ever_transplant = factor(ifelse(transplant == "None", "No", "Yes")))

#wbc distribution

overview_new %>%
  filter(ever_transplant == "Yes") %>%
  ggplot(aes(x= baseline_wbc)) +
  geom_histogram()

overview_new %>%
  filter(ever_transplant == "Yes") %>%
  group_by(trt_group) %>%
  mutate(N = n(),
         death = mean(inhosp_mortality28),
         se = sd(inhosp_mortality28)/sqrt(N), .groups = 'drop') %>%
  ggplot(aes(x = trt_group, y = death)) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin = death - se, ymax = death + se), width=0.2) +  # Add error bars
  labs(
    title = "Mortality rate by Treatment group among transplant patients",
    x = "Treatment group",
    y = "Mortality Rate"
  ) +
  theme_classic()
```
```{r}
#transplant modeling

# look only among transplant patients

 transplant_only <- overview_new %>%
  filter(ever_transplant == "Yes")

transplant_only_model <- interaction_model(transplant_only$inhosp_mortality28, transplant_only)

results <- model_results(transplant_only_model)
results

# add transplant to existing model

include_transplant <- glm(inhosp_mortality28 ~ trt_group +  baseline_wbc + baseline_wbc*trt_group + baseline_neutropenia + ever_transplant + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + location + preill_creat + priorhd, data = overview_new, family = "binomial")

new_results <- model_results(include_transplant)
new_results

```


## co-receipt of flagyl 

```{r}
new_smart_abx <- smart_abx %>%
  mutate(record_id = id) %>%
  left_join(acorn_emory_additional, by = "record_id")

new_smart_abx %>%
  group_by(trt_group, baseline_METRONIDAZOLE) %>%
  count()

new_smart_abx %>%
  group_by(trt_group, baseline_CLINDAMYCIN) %>%
  count()

new_smart_abx <- new_smart_abx %>%
  mutate(other_abx = ifelse(baseline_CLINDAMYCIN == "Yes" | baseline_METRONIDAZOLE == "Yes", "Yes", "No"))

new_smart_abx <- new_smart_abx %>%
  mutate(wbc_quartile = ntile(baseline_wbc, 4))

#add co-receipt to model

other_abx_model <- glm(inhosp_mortality28.x ~ trt_group +  baseline_wbc + baseline_wbc*trt_group + other_abx + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + location + preill_creat + priorhd, data = new_smart_abx, family = "binomial")

other_abx_results <- model_results(other_abx_model)
other_abx_results


#filter for people who did not also get anaerobes

no_other_abx <- new_smart_abx %>%
  filter(other_abx == "No")

no_other_abx_model <- glm(inhosp_mortality28.x ~ trt_group +  baseline_wbc + baseline_wbc*trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + location + preill_creat + priorhd, data = no_other_abx, family = "binomial")

no_other_abx_results <- model_results(other_abx_model)
no_other_abx_results

# look only in people who also got anaerobes

all_other_abx <- new_smart_abx %>%
  filter(other_abx == "Yes" | trt_group == "zosyn")

all_abx_model <- glm(inhosp_mortality28.x ~ trt_group +  baseline_wbc + baseline_wbc*trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + location + preill_creat + priorhd, data = all_other_abx, family = "binomial")

all_abx_results <- model_results(all_abx_model)
all_abx_results

all_abx_no_int <- glm(inhosp_mortality28.x ~ trt_group +  baseline_wbc + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + location + preill_creat + priorhd, data = all_other_abx, family = "binomial")

lrtest(all_abx_model, all_abx_no_int)

```

## OR data with no other antibiotics

```{r}
no_OR_data <- no_other_abx %>%
  group_by(wbc_epic) %>%
  nest() %>%
  mutate(
    fit = map(data, function(df) glm(inhosp_mortality28.x ~ trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + preill_creat + priorhd, data = df, family = "binomial")),
    output = map(fit, tidy),
    ci = map(fit, confint)) %>%
  unnest(c(output, ci)) %>%
  filter(term == "trt_groupzosyn") %>% 
  mutate(
    OR = exp(estimate),
    LL = exp(ci[,1]),
    UL = exp(ci[2])
  ) %>% 
  select(wbc_epic, OR, LL, UL)  %>%
  arrange(wbc_epic) 

no_OR_data$` ` <- paste(rep(" ", 35), collapse = " ")


# Create a confidence interval column to display
no_OR_data$`OR (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     no_OR_data$OR, no_OR_data$LL, no_OR_data$UL)


tm <- forest_theme(core=list(fg_params=list(hjust=01, x=0.9)), rowhead=list(fg_params=list(hjust=0.5, x=0.5)), padding = unit(c(50, 50), "mm"))

no_OR_forest <- forest(no_OR_data[, c(1, 5:6)],
            est = no_OR_data$OR,
            lower = no_OR_data$LL, 
            upper = no_OR_data$UL,
            ci_column = 2,
            ref_line = 1,
            arrow_lab = c("Favors piperacillin/tazobactam", "Favors Cefepime"),
            xlim = c(0, 3),
            ticks_at = c(0.5, 1, 1.5, 2, 2.5),
            title = "Mortality with Epic Cutoff with no Other Antibiotics", 
            theme = tm)
no_OR_forest
```

## look at culture data

cultures of interest are urine, sputum, blood, and other. unfortunately don't have additional data on what the other cultures were. All columns have potentially more than one value

Things to think about: 
- what to consider as primary source if more than one positive (esp if one is blood and same organim in urine or other source)
- just because culture positive does that mean that is the confirmed infectious source (i.e. does the organism matter)?
- 

```{r}

micro_cols <- c("sputum_cx_14day", "urine_cx_14day", "blood_cx_14day", "other_cx_14day")

#determine if all values are NA or no growth

all_neg <- function(x) {
  ifelse(is.na(x), TRUE,
         sapply(strsplit(x, ","), function(vals) {
           all(trimws(vals) %in% c("No growth", NA))
         }))
}

# identify source - blood, sputum, urine, other
new_smart_abx <- new_smart_abx %>%
  mutate(
    infectious_source = case_when(
      if_all(all_of(micro_cols), all_neg) ~ "Culture Negative",
      !all_neg(urine_cx_14day) & if_all(c(sputum_cx_14day, blood_cx_14day, other_cx_14day), all_neg) ~ "Urine",
      !all_neg(blood_cx_14day) & if_all(c(sputum_cx_14day, urine_cx_14day, other_cx_14day), all_neg) ~ "Blood",
      !all_neg(sputum_cx_14day) & if_all(c(urine_cx_14day, blood_cx_14day, other_cx_14day), all_neg) ~ "Sputum",
      !all_neg(other_cx_14day) & if_all(c(sputum_cx_14day, urine_cx_14day, blood_cx_14day), all_neg) ~ "Other",
      TRUE ~ "Multiple"   # fallback if >1 source is positive
    )
  )

new_smart_abx %>%
  group_by(infectious_source) %>%
  count()

new_smart_abx %>%
  filter(infectious_source == "Multiple") %>%
  select(micro_cols)
  head()
  
  
#identify organism

  bug_dict <- tribble(
  ~pattern,                 ~category,
  "^E\\. coli$",            "Gram-negative",
  "Klebsiella",             "Gram-negative",
  "Pseudomonas",            "Gram-negative",
  "Enterobacter",           "Gram-negative",
  "Staphylococcus",         "Gram-positive",   # catches aureus, epidermidis, etc.
  "Streptococcus",          "Gram-positive",   # catches pneumoniae, viridans, etc.
  "Enterococcus",           "Gram-positive",
  "Candida",                "Fungi",
  "Aspergillus",            "Fungi",
  "Acinetobactor",          "Gram-negative"

)


  # Split multiple values and clean
split_bugs <- function(x) {
  if (is.na(x)) return(NA_character_)
  bugs <- trimws(unlist(strsplit(x, ",")))
  if (all(bugs %in% c("No growth", NA))) return(NA_character_)
  return(bugs)
}

# Map bug names to categories
classify_bug <- function(bug, dict) {
  match <- dict %>%
    filter(grepl(pattern, bug, ignore.case = TRUE)) %>%
    slice_head(n = 1)
  
  if (nrow(match) == 0) {
    return("Unknown")
  } else {
    return(match$category)
  }
}

#vectorized classify bug to run faster

# Expand each regex into a column match
classify_bugs_fast <- function(bugs, dict) {
  # For each bug, find first matching pattern
  out <- rep("Unknown", length(bugs))
  for (i in seq_len(nrow(dict))) {
    hits <- grepl(dict$pattern[i], bugs, ignore.case = TRUE)
    out[hits & out == "Unknown"] <- dict$category[i]
  }
  return(out)
}


new_smart_abx <- new_smart_abx %>%
  rowwise() %>%
  mutate(
    # assign source
    infectious_source = case_when(
      if_all(all_of(micro_cols), ~ all_neg(.x)) ~ "Culture Negative",
      !all_neg(urine_cx_14day) & if_all(c(sputum_cx_14day, blood_cx_14day, other_cx_14day), all_neg) ~ "Urine",
      !all_neg(blood_cx_14day) & if_all(c(sputum_cx_14day, urine_cx_14day, other_cx_14day), all_neg) ~ "Blood",
      !all_neg(sputum_cx_14day) & if_all(c(urine_cx_14day, blood_cx_14day, other_cx_14day), all_neg) ~ "Sputum",
      !all_neg(other_cx_14day) & if_all(c(sputum_cx_14day, urine_cx_14day, blood_cx_14day), all_neg) ~ "Other",
      TRUE ~ "Multiple"
    ),
    # extract bugs from urine column (do same for others if you like)
    urine_bugs = list(split_bugs(urine_cx_14day)),
    blood_bugs = list(split_bugs(blood_cx_14day)),
    sputum_bugs = list(split_bugs(sputum_cx_14day)),
    other_bugs  = list(split_bugs(other_cx_14day))
  ) %>%
  ungroup()

# look at infectious source

new_smart_abx %>%
  group_by(trt_group, infectious_source) %>%
  count()

#does culture negative go down if you look only in sepsis?

new_smart_abx %>%
  filter(sepsis == "Yes") %>%
  group_by(trt_group, infectious_source) %>%
  count()

#classify bugs, slow
bugs_long <- new_smart_abx %>%
  select(id, infectious_source, inhosp_mortality28.x, trt_group, baseline_wbc, age, sex,  baseline_sofa, baseline_vent, baseline_anyvasopressor, location, preill_creat, priorhd,
         urine_bugs, blood_bugs, sputum_bugs, other_bugs) %>%
  pivot_longer(cols = ends_with("_bugs"), names_to = "source", values_to = "bugs") %>%
  unnest(bugs) %>%
  mutate(
    category = sapply(bugs, classify_bug, dict = bug_dict)
  )

#classify bugs, fast
bugs_long <- new_smart_abx %>%
  select(id, infectious_source, inhosp_mortality28.x, trt_group, baseline_wbc, age, sex,  baseline_sofa, baseline_vent, baseline_anyvasopressor, location, preill_creat, priorhd,
         urine_bugs, blood_bugs, sputum_bugs, other_bugs) %>%
  tidyr::pivot_longer(cols = ends_with("_bugs"), names_to = "source", values_to = "bugs") %>%
  unnest(bugs) %>%
  mutate(category = classify_bugs_fast(bugs, bug_dict))

# unknown bugs for review

unknown_bugs <- bugs_long %>%
  filter(category == "Unknown") %>%
  distinct(bugs) %>%
  arrange(bugs)


# This gives you a tidy list of bugs to review and add to your dictionary
unknown_bugs



# look at microbiologic causes

bugs_long %>%
  group_by(category,trt_group) %>%
  count()

# add rows for missed important bacteria

bug_dict <- bug_dict %>%
  add_row(pattern = "Acinetobacter", category = "Gram-negative") %>%
  add_row(pattern = "Proteus", category = "Gram-negative")


```

### modeling with micro cause

```{r}
micro_source_model <- glm(inhosp_mortality28.x ~ trt_group +  baseline_wbc + baseline_wbc*trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + factor(infectious_source) + location + preill_creat + priorhd, data = new_smart_abx, family = "binomial")

micro_source_results <- model_results(micro_source_model)
micro_source_results

micro_source_no_int_model <- glm(inhosp_mortality28.x ~ trt_group +  baseline_wbc + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + factor(infectious_source) + location + preill_creat + priorhd, data = new_smart_abx, family = "binomial")

micro_source_no_int_results <- model_results(micro_source_no_int_model)
micro_source_no_int_results

lrtest(micro_source_model, micro_source_no_int_model)

# modeling with bug type

bug_model <- glm(inhosp_mortality28.x ~ trt_group +  baseline_wbc + baseline_wbc*trt_group + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + factor(category) + location + preill_creat + priorhd, data = bugs_long, family = "binomial")

bug_results <- model_results(bug_model)
bug_results

bug_no_int_model <- glm(inhosp_mortality28.x ~ trt_group +  baseline_wbc + age + sex+  baseline_sofa + baseline_vent + baseline_anyvasopressor + factor(category) + location + preill_creat + priorhd, data = bugs_long, family = "binomial")

lrtest(bug_model, bug_no_int_model)

```


## code to create survival plot


```{r}
library(survival)
install.packages("survminer")
library(survminer)

# Convert inhosp_mortality28 to binary numeric
new_smart_abx <- new_smart_abx %>%
  mutate(event = as.numeric(inhosp_mortality28.x == "Yes"),
         WBC_Quartile = factor(wbc_quartile, 
                               levels = c(1, 2, 3, 4), 
                               ordered = TRUE, 
                               labels = c("<7", "7-11", "11-16", ">16")))   

# Create a survival object
surv_obj <- Surv(time = new_smart_abx$hours_start_death, event = new_smart_abx$event)

# Fit the survival curve by trt_group
fit <- survfit(surv_obj ~ trt_group, data = new_smart_abx)

# Plot the curve
ggsurvplot(fit, data = new_smart_abx, pval = TRUE, risk.table = TRUE,
           legend.title = "Treatment Group",
           xlab = "Days", ylab = "Survival Probability")

install.packages("ggsurvfit")
library(ggsurvfit)

fit1 <- survfit2(Surv(hours_start_death, inhosp_mortality28.x) ~ trt_group, data = new_smart_abx) 

ggsurvplot_facet(fit1, new_smart_abx, facet.by = "wbc_epic")

ggsurv

#survival curve by WBC quartile

fit2 <- survfit2(Surv(hours_start_death, inhosp_mortality28.x) ~ trt_group, data = new_smart_abx) 


survival_curve <- ggsurvplot_facet(fit2, new_smart_abx, facet.by = "WBC_Quartile") +
  theme_minimal()

#new fit2

fit2 <- survfit2(Surv(hours_start_death, event) ~ trt_group, data = new_smart_abx)
survival_curve <- ggsurvplot_facet(fit2, new_smart_abx, facet.by = "WBC_Quartile") +
  theme_minimal()

survival_curve <- ggsurvplot(fit2, new_smart_abx)

fit3 <- survfit2(Surv(hours_start_death, inhosp_mortality28.x) ~ trt_group + WBC_Quartile, data = new_smart_abx)
survival_curve2 <- ggsurvplot_facet(fit3, new_smart_abx, facet.by = "WBC_Quartile") +
  theme_minimal()


```



### calculate absolute risk difference and predicted probability at different wbc levels
```{r}
smart_abx <- smart_abx %>%
  mutate(across(where(is.character), as.factor),
         baseline_anyvasopressor = as.factor(baseline_anyvasopressor))

#model of interest

moda <- glm(inhosp_mortality28 ~ trt_group + baseline_wbc + baseline_wbc*trt_group + age + sex+ baseline_sofa + baseline_vent + baseline_anyvasopressor + source_of_infection + location + preill_creat + priorhd, data = smart_abx, family = "binomial")

# Example WBC levels to evaluate
wbc_levels <- c(7, 11, 16)

# Build a "newdata" grid for prediction
pred_data <- expand.grid(
  baseline_wbc = wbc_levels,
  trt_group = levels(smart_abx$trt_group),
  age = mean(smart_abx$age, na.rm = TRUE),
  sex = "Male",   # choose reference or common value
  baseline_sofa = mean(smart_abx$baseline_sofa, na.rm = TRUE),
  baseline_vent = "No",   # set to reference/typical
  baseline_anyvasopressor = "1",
  source_of_infection = "Lung",
  location = "Medical ICU",
  preill_creat = mean(smart_abx$preill_creat, na.rm = TRUE),
  priorhd = "No"
)

#obtain predicted probabilities
pred_data$pred_prob <- predict(moda, newdata = pred_data, type = "response")

#calculate risk differences

risk_diff <- pred_data %>%
  group_by(baseline_wbc) %>%
  summarise(
    abs_risk_diff = diff(pred_prob),  # assumes exactly 2 groups
    .groups = "drop"
)

# predicted probability and risk difference in one place

pred_table <- pred_grid %>%
  tidyr::pivot_wider(names_from = trt_group, values_from = pred_prob) %>%
  mutate(abs_risk_diff = !!sym(levels(smart_abx$trt_group)[2]) - 
                        !!sym(levels(smart_abx$trt_group)[1])) %>%
  select(baseline_wbc, cefepime, zosyn, abs_risk_diff) %>%
  filter(baseline_wbc %in% c(7, 11, 16))

pred_table
```


```{r}
#graph of predicted probabilities

ggplot(pred_data, aes(x = baseline_wbc, y = pred_prob, color = trt_group)) +
  geom_line() +
  geom_point() +
  labs(y = "Predicted Mortality Probability",
       x = "Baseline WBC",
       title = "Predicted Mortality by WBC and Treatment")
```
### smooth curve of predicted probability by wbc 

```{r}
# Range of WBC values (e.g. 1 to 30)
wbc_seq <- seq(1, 35, by = 0.5)

pred_grid <- expand.grid(
  baseline_wbc = wbc_seq,
  trt_group = levels(smart_abx$trt_group)
)

# Use a template row to preserve factor levels
template <- smart_abx[1, ]

pred_grid$age <- mean(smart_abx$age, na.rm = TRUE)
pred_grid$sex <- template$sex
pred_grid$baseline_sofa <- mean(smart_abx$baseline_sofa, na.rm = TRUE)
pred_grid$baseline_vent <- template$baseline_vent
pred_grid$baseline_anyvasopressor <- template$baseline_anyvasopressor
pred_grid$source_of_infection <- template$source_of_infection
pred_grid$location <- template$location
pred_grid$preill_creat <- mean(smart_abx$preill_creat, na.rm = TRUE)
pred_grid$priorhd <- template$priorhd

#predict probabilities

pred_grid$pred_prob <- predict(moda, newdata = pred_grid, type = "response")

# plot curve

ggplot(pred_grid, aes(x = baseline_wbc, y = pred_prob, color = trt_group)) +
  geom_line(size = 1.2) +
  labs(
    x = "Baseline WBC",
    y = "Predicted 28-day Mortality Probability",
    title = "Predicted Mortality by WBC and Treatment"
  ) +
  theme_minimal()
```
### table one with neutropenia and transplant

```{r}
##creating Table 1
#characteristics to capture: age, sex, race, smoking, baseline creatinine, Charlson, baseline qSOFA, infection source

#table 1 with patient level stuff
myVars <- c("trt_group", "vancomycin_day0","received_vancomycin_14", "age",  "bmi",  "baseline_sofa", "daysaccurate_timezero_firstanyantibiotics", "CHARLSON", "sex", "race.eth", "baseline_vent", "prevalent_aki.factor", "sepsis", "source_of_infection", "baseline_anyvasopressor", "inhosp_mortality28", "MAKE14", "comagdelcfreedays14", "baseline_coma_delirium", "baseline_neutropenia", "transplant")

catVars <- c("trt_group", "vancomycin_day0","received_vancomycin_14","sex", "race.eth", "baseline_vent", "prevalent_aki.factor", "sepsis", "source_of_infection", "baseline_anyvasopressor", "inhosp_mortality28", "baseline_neutropenia", "transplant")

table1 <- CreateTableOne(vars = myVars, data = overview_new, factorVars = catVars, strata = "trt_group")
table1

table1 <- print(table1, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
## Save to a CSV file
write.csv(table1, file = "myTable.csv")

#repeat, stratified by treatment group
table1_trt_group <- CreateTableOne(vars = myVars, overview_new, factorVars = catVars, strata = c("trt_group"))
table1_trt_group

#stratified by wbc quartile
table1_quartile <- CreateTableOne(vars = myVars, overview_new, factorVars = catVars, strata = c("wbc_quartile"))
table1_quartile
```
### wbc distribution based on septic vs not

```{r}
new_smart_abx %>%
  group_by(sepsis) %>%
  summarise(mean = mean(baseline_wbc),
            median = median(baseline_wbc),
            range = range(baseline_wbc),
            min = min(baseline_wbc),
            max = max(baseline_wbc))
```


```{r}
#data for survival plot

new_smart_abx <- new_smart_abx %>%
  mutate(event = as.numeric(inhosp_mortality28.x == "Yes"),
         WBC_Quartile = factor(wbc_quartile, 
                               levels = c(1, 2, 3, 4), 
                               ordered = TRUE, 
                               labels = c("<7", "7-11", "11-16", ">16")))   

# Create a survival object
surv_obj <- Surv(time = new_smart_abx$hours_start_death, event = new_smart_abx$event)

surv_plot_data <- new_smart_abx %>%
  select(id, wbc_quartile, wbc_epic, inhosp_mortality28.x, hours_start_death, trt_group)

# Export the data frame to a CSV file
write_csv(surv_plot_data, "surv_plot_data.csv")
```

